---
title: "MonteCarloMed"
author: "zzh"
date: "7/9/2019"
output: 
    prettydoc::html_pretty:
      highlight: github
      theme: architect
      toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This is my first try to apply Monte Carlo method to clinical medicine to choose an treatment strategy that is best suitable for a condition. Monte Carlo method is thought to be the best suitable model for solving clinical problems becase the sequence is a patient's trajectory and is avaiable during data collection. The full sequence is also avaiable at the end of a study, which is unlike some other situations such as blackjack and cliff walking games, in which the generation of sequence is cheap and can be on-line.  

# DataSimulation
```{r DataGeneration}
N=3000
set.seed(123)
id <- sample(x=1:500,size = N,replace = T)
map <- rnorm(N,mean = 60,sd=10)
lac <- abs(rnorm(N,mean = 3,sd = 1))
hct <- rnorm(N,mean = 30,sd=10)
actions <- 1:10
dt <- data.frame(id=id,map=map,lac=lac,hct=hct,actions=actions)
dt <- dt[order(dt$id),]
head(dt)
library(plyr)
dt <- ddply(dt,.(id),function(dd){
  dd$mort <- ifelse(1:nrow(dd)==nrow(dd),rbinom(1,1,0.4),0)
  return(dd)
})
```

#Use linear approximation; may explode
```{r LinearApproximation}
Weight<-matrix(rep(0,5),ncol = 1)
alpha = 0.0005
for (episode in 1:10) {
  dtsample <- dt[dt$id==episode,]
  Nsteps <- nrow(dtsample)
  G <- -sum(dtsample[,"mort"])
  V <- dtsample[,c("map","lac","hct","actions")]
  V$intercept <- 1
  for (step in 1:Nsteps) {
    Weight <- Weight + as.vector(alpha*(G - as.numeric(V[step,]) %*% Weight))*t(V[step,])
    
  }
  if(sum(is.na(Weight))>0){
      break;
    }
}
Weight
#the result cannot converge due to randomly generated samples
```

#Using caret package to train the model
```{r CaretPackage}
#return for each episode is the same -1 for mort=1; 1 for mort = 0 (Not an idea choice)
dt <- ddply(dt,.(id),function(dd){
  dd$G <- ifelse(sum(dd$mort)==0,1,-1)
  return(dd)
})
#discounted return gamma=0.95
Gamma <- 0.95
dt <- ddply(dt,.(id),function(dd){
  steps <- nrow(dd)
  for (i in 1:steps) {
    if(sum(dd$mort)==0){
      dd$G[i] <- Gamma^(steps-i)
    }else{
      dd$G[i] <- -Gamma^(steps-i)
    }
  }
  return(dd)
})
```

#Model taining 
```{r ModelTraining, results="hide"}
library(caret)
#model training
NNGrid <-  expand.grid(size = 3:10, 
                        decay = seq(0,0.1,by=0.02))
mod <- train(G~map+lac+hct+actions,data = dt,
             method="nnet",
             tuneGrid = NNGrid,
             verbose = FALSE)
```

#Visulization
```{r Visulization}
plot(mod)
#to plot how to choose action based on the model
index = 20 # a stuation like patient 20; 
dtPred <- dt[rep(index,10),]
dtPred$actions <- 1:10 # vary the actions to see which can achieve the largest action value
action_V <- predict.train(mod,newdata = dtPred)
plot(dtPred$actions,action_V)
```